miniKanren with fair search strategies

### 2019-04-30 Paper/paper.tex
lots of update

%TODO our new searches also change the behavior of conj
The syntax of a programming language should reflect its semantics. When using a
disjunction operator in relational programming, a programmer would expect all
clauses of this disjunct to share the same chance of being explored, as
these clauses are written in parallel. The existing multi-arity disjunctive
operator in miniKanren, however, prioritize its clauses by the order of which
these clauses are written down. We have devised two new search strategies that
allocate computational effort more fairly in all clauses.

@@

miniKanren programs, especially relational interpreters, have been proven to be useful in solving many problems \citep{byrd2017unified}. A subtlety in writing relational programs involving a large \conde expression, such as interpreters, is that the order of \conde clauses can affect the speed considerably. This is due to the biased allocation of computational resource.
Interleaving DFS, the search strategy of \conde in miniKanren~\citep{friedman_reasoned_2018}, allocates half resource to the first \conde clause, a quarter to the second, an eighth to the third, and so on. Under the hood, it is the unfairness of \texttt{disj}.

We address the unfairness of \texttt{disj} in this work. We propose two new search strategies, balanced interleaving DFS (biDFS) and breadth-first search (BFS). biDFS is almost fair -- the maximal ratio of resource among disjunctive goals is bounded by a constant factor. BFS is completely fair. We prove that our BFS is equivalent to the BFS proposed by Seres et al~\citep{seres1999algebra}. We also observe how new search strategies affect the efficiency and the answer order of known miniKanren programs.

@@

A search strategy is \emph{fair} if answers of lower costs always come first. 
The \emph{cost} of an answer is the number of relation applications. Now we 
illustrate the costs of answers by running a miniKanren relation. 
Fig.~\ref{def-repeato} defines a relation \texttt{repeato} which relates a term 
\texttt{x} with a list whose elements are all \texttt{x}s.

%TODO compare our definition with Silvija's, Oleg's, and Edward's.

\begin{figure}
  \lstinputlisting{Figures/defrel-repeato.rkt}
  \caption{\texttt{repeato}}
  \label{def-repeato}
\end{figure}

Consider the following \texttt{run} of \texttt{repeato}.
\begin{center}
  \begin{tabular}{c}
  \lstinputlisting{Figures/run-repeato-1.rkt}
   \end{tabular}
\end{center}

The above \texttt{run} generates 4 answers. All are lists of \texttt{*}s.
The order of the answers reflects the order miniKanren discovers them:
the first answer in the list is first discovered. This order is not surprising:
to generate the first answer, \texttt{'()}, miniKanren needs to apply
\texttt{repeato} only once and the later answers need more relation
applications. In this example, the cost of each answer is the same as
one more than the number of \texttt{*}s: the cost of \texttt{'()} is 1,
the cost of \texttt{'(*)} is 2, and so on.

% A query result (list of answers) is in the \emph{cost-respecting} order if no 
% answer occurs before another answer of a lower cost. In the above example, the 
% result is cost-respecting. The iDFS search, however, does not generate 
% cost-respecting answers in general. As an example, consider the following 
% \texttt{run} of \texttt{repeato}.

In the above example, every search strategy looks fair. However, the following 
example exposes that iDFS is not fair.

\begin{center}
  \begin{tabular}{c}
   \lstinputlisting{Figures/run-conde-repeato-iDFS.rkt}
   \end{tabular}
\end{center}

With iDFS, \texttt{'(a a)} occurs before \texttt{'(b)} while \texttt{'(a a)} is 
associated with a higher cost. iDFS strategy is the cause since it prioritizes 
the first \conde case considerably. When every \conde case 
are equally productive, the iDFS strategy takes $1/2^{i}$ answers from the 
$i$-th case, except the last case, which share the same portion as the second 
last one. In contrast, the same run with BFS produces answers in an expected 
order.

\begin{center}
  \begin{tabular}{c}
   \lstinputlisting{Figures/run-conde-repeato-BFS.rkt}
   \end{tabular}
\end{center}

Running the same query with biDFS results in yet another answer list. biDFS 
essentially organize disjunctive goals into a balanced tree. There is no way to 
build a balanced and complete tree of size 3, so one clause is allocated more 
resource than the other two.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/run-conde-repeato-biDFS.rkt}
	\end{tabular} 
\end{center}

If one insert a \texttt{(nevero)} as the forth clause, this run would results 
in the same answer list as the run with BFS. However, just making every \conde 
has $2^n$ clauses cannot turn biDFS to BFS.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/run-conde-repeato-biDFS-1.rkt}
	\end{tabular}
\end{center}

@@

Our first solution, balanced interleaving DFS (biDFS), like iDFS, is not fair . However, it is less sensitive to goal order in disjunct and is as efficient as iDFS.

The reason why iDFS's \texttt{disj} prioritizes its left goals considerably is that the \texttt{disj} applys \texttt{disj2} right associatively, and that 
\texttt{disj2} allocates resource evenly to its two sub-goals. If a disjunct 
is viewed as a binary tree where \texttt{disj2}s are nodes and sub-goals are 
leaves, the deeper a leaf locates, the lower resource it is shared. In iDFS, the tree is in one of the most unbalanced forms.

The key idea of biDFS is to make the tree balanced. Fig.~\ref{balanced-disj} shows the difference between iDFS and biDFS. We introduce a function \texttt{disj*} and its helper \texttt{split}, and change the \texttt{disj} macro to call \texttt{disj*} immediately. \texttt{disj*} essentially construct a balanced \texttt{disj2} tree. The \texttt{split} helper splits elements of \texttt{ls} into two lists of roughly the same length, then apply \texttt{k} to the two sub-lists.

@@

@@

In this section, we describe our BFS and compare it with the one from Seres et 
al \citep{seres1999algebra}. The first subsection is devoted to introducing BFS to miniKanren. This subsection results in a new version of miniKanren, \texttt{mk-1} (we call the original version \texttt{mk-0} for short). The second subsection describes the equivalent between \texttt{mk-1} and Silvija's BFS. In the third and last subsection, we optimize \texttt{mk-1} with the help of a queue, which results in \texttt{mk-2}, the final BFS version.

\subsection{change search strategy from iDFS to BFS}

In both \texttt{mk-0} and \texttt{mk-1}, search spaces are represented by 
streams of answers. Thunk streams in \texttt{mk-0} denote delayed 
computation, however, they do not necessarily mean an increment in cost. We use 
the same kind of stream in \texttt{mk-1} but only put thunk at those places 
where an increment in cost happens.

For convenience, we call the \texttt{car}s of a stream as its \emph{mature} 
part, and its last \texttt{cdr} as its \emph{immature} part. When the stream is 
definitely finite, its immature part is an empty list, otherwise, it is a 
thunk. We sometimes say a stream is immature to mean its mature part is empty.

Streams denote cost correctly when they are constructed by \texttt{==}, 
\texttt{succeed}, and \texttt{fail}. However, the \texttt{mk-0} version of 
\texttt{append-inf} (Fig.~\ref{append-inf-0}) breaks the rule when its first input stream, \texttt{s-inf}, has a non-trivial immature part. In this case, the resulting mature part contains only the mature part of \texttt{s-inf}. If we want to describe the cost information with thunks, the resulting mature part should also contain the mature part of \texttt{t-inf}.

\begin{figure}
	 \lstinputlisting{Figures/append-inf-0.rkt}	
	 \caption{\texttt{append-inf} in \texttt{mk-0}}
	 \label{append-inf-0}
\end{figure}

The \texttt{mk-1} version of \texttt{append-inf} (Fig.~\ref{append-inf-1}) 
gain fairness by combining the mature parts in the fashion of \texttt{append}. 
This \texttt{append-inf} calls its helper immediately, with the first argument, 
\texttt{s?}, set to \texttt{\#{}t}, which indicates whether \texttt{s-inf} and \texttt{t-inf} haven't been swapped in the driver. \texttt{s-inf} and \texttt{t-inf} are swapped in the third \texttt{cond} clause, where \texttt{s?} is flipped accordingly.

\begin{figure}
	 \lstinputlisting{Figures/append-inf-1.rkt}	
	 \caption{\texttt{append-inf} in \texttt{mk-1}}
	 \label{append-inf-1}
\end{figure}

\texttt{mk-1} is inefficient in two aspects. \texttt{append-inf} need to copy 
all \texttt{cons} cells of \emph{both} input streams when the first stream is 
has a non-trivial immature part. Besides, \texttt{mk-1} computes answers of the same cost at once, even when only a small portion is queried. We solve the two problems in the next subsections.

\subsection{compare our BFS with Seres's}

;; under construction

\subsection{optimize breadth-first search}

We avoid generating same-cost answers at once by expressing BFS with a queue, 
whose elements are thunks that return a new stream. Every \texttt{mk-1} stream has zero or one thunk, so it is uninteresting to manage them with the queue. Therefore, we change the representation of immature parts from thunks to 
thunk lists. As a side effect, it is no longer convenient to combine the mature and immature part with \texttt{append}, which would mix answers and thunks in the same list. We choose \texttt{cons} as an alternative to \texttt{append}.

After applying these two changes, stream representation becomes more 
complicated, which motivates us to set up an interface between stream 
and the rest of miniKanren. Listed in Fig.~\ref{interface} are all functions 
being aware of the stream representation, except \texttt{take-inf} and its helper function (they are explained later).

The first three functions are constructors: \texttt{empty-inf} makes an 
empty stream; \texttt{unit} makes a stream with one mature 
solution; \texttt{step} makes a stream with one thunk.

\texttt{append-inf} combines each sub-parts with \texttt{append}.

\texttt{append-map-inf} ...

The next four functions are only depended on by \texttt{ifte} and \texttt{once}. \texttt{null-inf?} checks whether a stream is exhausted. \texttt{mature-inf?} checks whether a stream has some mature 
solutions. \texttt{car-inf} takes the first solution out of a mature stream. 
\texttt{cdr-inf} drops the first solution of a mature stream. Finally, 
\texttt{force-inf} forces an immature stream to do more computation.

\texttt{unit}, \texttt{append-map-inf}, \texttt{empty-inf} and 
\texttt{append-inf} form a \emph{MonadPlus}, where they correspond to 
\texttt{unit}, \texttt{bind}, \texttt{mzero}, and \texttt{mplus} respectively.

\begin{figure}
	 \lstinputlisting{Figures/interface.rkt}	
	 \caption{Functions being aware of stream representation}
	 \label{interface}
\end{figure}

The last interesting function is \texttt{take-inf} (Fig.~\ref{take-inf-3-1}). 
Its first parameter, \texttt{vs}, is a list of solutions. The next two parameters, \texttt{P} and \texttt{Q}, together represents a queue. 
The first two \texttt{cond} lines are very similar to their counterparts in mk-0 and mk-1. The third line runs when both the answer list \texttt{vs} and the queue are empty, which means we have found all the answers. The fourth line re-shape the queue. The last line invokes the first thunk in the queue and use the mature part of the resulting stream, \texttt{s-inf}, as the new \texttt{vs}, and enqueuing \texttt{s-inf}'s thunks.

\begin{figure}
	 \lstinputlisting{Figures/take-inf-3-1.rkt}	
	 \caption{\texttt{take-inf} in \texttt{mk-3-1}}
	 \label{take-inf-3-1}
\end{figure}

@@

%TODO evalo (saying I love you), appendo, reverseo
;; Kuang-Chen and Weixi plan to put '(I love you), quines, appendo, and reverso here

@@

;; under construction

Edward points out a disjunct would be `fair' if its tree representation is balanced and full \citep{yang2010adventures}.

Silvija et al \citep{seres1999algebra} also describe a breadth-first search strategy. We proof their BFS is equivalent to ours. However, ours looks simpler and runs about twice faster in comparison with a straightforward translation of their Haskell code.

@@

\subsection{others}
%TODO contributions
%TODO future works

We devise a new search strategy, balanced interleaving DFS. The key idea is to 
make disjunct trees balanced. Changing the search strategy from iDFS to biDFS is not hard: 2 new functions and 1 modified macro.

We also devise breadth-first search, whose intuition is similar to Seres's BFS.
And we have proved their equivalence. We optimize our BFS with a queue.

>>>>

%TODO rewrite abstract
%The syntax of a programming language should reflect its semantics. When using a disjunctive operator in relational programming, a programmer would expect all sub-goals of this disjunct to share the same chance of being explored, as these clauses are written in parallel. The existing multi-arity disjunctive operator in miniKanren, however, prioritize its clauses by the order of which these clauses are written down. We have devised two new search strategies that allocate computational effort more fairly in all clauses.

@@

miniKanren programs, especially relational interpreters, have been proven to be useful in solving many problems \citep{byrd2017unified}. A subtlety in writing relational programs having large \conde expressions, such as interpreters, is that the order of \conde clauses can affect the speed considerably. When a \conde expression is large enough, the left clauses consume almost all the resource and the right ones are hardly explored. The unfair \disj{} of the current search strategy, interleaving DFS, is the cause. Under the hood, \conde uses \texttt{conj} to create a goal for each clause, and \texttt{disj} to combine these goals to one. The current \texttt{disj} allocates half resource to its first goal, then allocates the other half to the rest similarly, except for the last clause which takes all the resource.

Being aware of \texttt{disj} fairness, we also investigate \texttt{conj} fairness.

We propose three new search strategies, balanced interleaving DFS (biDFS), fair DFS (fDFS), and BFS. They have different characteristics of fairness (Table.~\ref{fairness}). We prove that our BFS and the BFS proposed by Seres et al~\citep{seres1999algebra} produce the same result when queried. But our code is shorter and runs faster. We also compare the efficiency of these new search strategies with the existing ones.

\begin{table}
	\begin{tabular}{|c|c|c|c|c|}
		\hline 
		fairness & iDFS & biDFS & fDFS & BFS \\ 
		\hline 
		disj & unfair & almost-fair & fair & fair \\ 
		\hline 
		conj & unfair & unfair & unfair & fair \\ 
		\hline 
	\end{tabular} 
	\caption{fairness of search strategies}
	\label{fairness}
\end{table}

@@

We demonstrate the aspects (\disj{} or \conj{}) and levels (unfair, almost-fair, or fair) of fairness by running queries about \texttt{repeato}, a relational definition that relates a term \texttt{x} with a non-empty list whose elements are \texttt{x} (Fig.~\ref{repeato}).

\begin{figure}
	\lstinputlisting{Figures/repeato.rkt}
	\caption{\texttt{repeato} and an example run}
	\label{repeato}
\end{figure}

\subsection{fair \texttt{disj}}

In the following program, the three \conde clauses differ in a trivial way. So we expect lists of each sort constitute $1/4$ of the answer list. However, iDFS, the current search strategy, gives us much more lists of \texttt{a} than other sorts of lists. And some sorts (e.g. lists of \texttt{c}) are hardly found. The situation would be worse if we add more \conde clauses.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/repeato-disj-iDFS.rkt}
	\end{tabular}
\end{center}

On the contrary, search strategies with fair \disj{} (e.g. fDFS and BFS) give a nice answer list.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/repeato-disj-fair.rkt}
	\end{tabular}
\end{center}

Search strategies with almost-fair \disj{} (e.g. biDFS) give the same result in this case. However, as its name suggests, almost-fair strategies are not always fair. Fortunately, the maximal ratio of the allocated resource is bounded by a constant. Our biDFS is fair when the number of goals is a power of 2, otherwise, some goals are allocated twice more resource than the others. In following example, the clauses of \texttt{b}, \texttt{c}, and \texttt{d} are allocated more resource.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/repeato-disj-biDFS.rkt}
	\end{tabular}
\end{center}

\subsection{fair \texttt{conj}}

In the following program, the three \conde clauses differ in a trivial way. So we expect lists of each sort constitute  $1/4$ of the answer list. However, search strategies with unfair \conj{} gives us much more lists of \texttt{a} than other sorts of lists. And some sorts (e.g. lists of \texttt{c}) are hardly found. The situation would be worse if we add more \conde clauses. The result with biDFS, whose \conj{} is also unfair, is similar, but due to its different \disj{}, the position of \texttt{b} and \texttt{c} and swapped.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/repeato-conj-unfair.rkt}
	\end{tabular}
\end{center}

On the contrary, search strategies with fair \conj{} (e.g. BFS) give a nice answer list.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/repeato-conj-fair.rkt}
	\end{tabular}
\end{center}

A more interesting situation is when the first conjunctive goal produces infinite many states. Consider the following example, a naive specification of fair \conj{} might require search strategies to produce all sorts of singleton lists, but no longer lists, which makes the strategies \emph{incomplete}.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/repeato-conj-infinite-naive.rkt}
	\end{tabular}
\end{center}

Fairness is good but is not good enough to kick away completeness. A solution also taken in \citep{seres1999algebra} is bagging states and requiring that search spaces derived from states in the same bag are treated fairly. A natural way to bag states is by their costs. The \emph{cost} of a state is its depth in the search tree (i.e. the number of calls to relational definitions required to find them) \citep{seres1999algebra}. BFS is more than fair -- it produces answers in increasing order of cost. Running the same program gives an answer list sorted by answers' costs. In this case, the cost of an answer is equal to the length of the inner lists plus the length of the outer list.

\begin{center}
	\begin{tabular}{c}
		\lstinputlisting{Figures/repeato-conj-infinite-sof.rkt}
	\end{tabular}
\end{center}

@@

Balanced interleaving DFS (biDFS) has almost-fair \disj{} and unfair \conj{}. The implementation of biDFS differs from iDFS in the \disj{} macro. We list the new \disj{} with its helpers in Fig.~\ref{balanced-disj}. The first helper function, \texttt{split}, takes a list of goals \texttt{ls} and a procedure \texttt{k}, partitions \texttt{ls} into two sub-lists of roughly equal length, and returns the application of \texttt{k} to the two sub-lists. \texttt{disj*} takes a non-empty list of goals \texttt{gs} and returns a goal. With the help of \texttt{split}, it essentially constructs a \emph{balanced} binary tree where leaves are goals in \texttt{gs} and nodes are \texttt{disj2}, whence the name of this search strategy. In contrast, the \disj{} in iDFS essentially constructs the same sort of binary tree in one of the most unbalanced forms.

@@

\section{fair DFS}

Fair DFS (fDFS) has fair \disj{} and unfair \conj{}. The implementation of fDFS differs from iDFS's in \texttt{disj2} (Fig.~\ref{fDFS}). \texttt{disj2} is changed to call a new and fair version of \texttt{append-inf}. \texttt{append-inf/fair} immediately calls its helper, \texttt{append-inf/fair\^{}}, with the first argument, \texttt{s?}, set to \texttt{\#{}t}, which indicates that \texttt{s-inf} and \texttt{t-inf} haven't been swapped. The swapping happens in the third \texttt{cond} clause of the helper, where \texttt{s?} is changed accordingly. The first two \texttt{cond} clauses essentially copy the \texttt{car}s and stop recursion when one of the input spaces is obviously finite. The third clause, as we mentioned early, is just for swapping. When the fourth and last clause runs, we know that both \texttt{s-inf} and \texttt{t-inf} are ended with a thunk. In this case, a new thunk is constructed. The new thunk calls the driver recursively. Here changing the order of \texttt{t-inf} and \texttt{s-inf} won't hurt the fairness (though it will change the order of answers). We swapped them back so that answers are produced in a more natural order.

\begin{figure}
	\lstinputlisting{Figures/fDFS.rkt}
	\caption{How fDFS differs from iDFS}
	\label{fDFS}
\end{figure}

@@

Our BFS is fair in both \disj{} and \conj{}. Its implementation is based on fDFS (not iDFS). All we have to do is changing the use of \texttt{append-inf} in \texttt{append-map-inf} to \texttt{append-inf/fair}. After making this change, \texttt{append-inf} is useless.

The implementation can be improved in two aspects. First, as we mentioned in section 2.2, our BFS bag states by their cost. However, in this implementation, it is unclear where this information is recorded. Second, \texttt{append-inf/fair} is space inefficient. It makes $O(n+m)$ new \texttt{cons} cells every time, where $n$ and $m$ is the ``length'' of each input search space. We address these issues in the first subsection.

Both our BFS and Seres's BFS \citep{seres1999algebra} produce answers in increasing order of cost. So it is interesting to see if they are equivalent. We prove the equivalence in Coq. The details are in the second subsection.

% The first subsection describe an optimized BFS. The optimization makes the type of search space representation more clear and reduces time-complexity of some function. The second subsection compares the pure subset of the optimized BFS with the BFS found in \citep{seres1999algebra}. We prove the two BFSs return the same answer list.

\subsection{optimized BFS}

As we mentioned in section 2.2, our BFS bag states by their cost. The bagging information is recorded subtly -- the \texttt{car}s of a search space have cost 0 (they are in the same bag), and the costs of states in thunk are increased by one. It is even more difficult to see \texttt{append-inf/fair} and the modified \texttt{append-map-inf} respects the cost information. We make things clear by changing the type of search space, modify related function definitions, and introducing a few more functions.

The new type is a pair whose \texttt{car} is a list of state (the bag), and whose \texttt{cdr} is either a \texttt{\#{}f} or a thunk returning a search space. A falsy \texttt{cdr} indicates that the search space is obviously finite.

Both modified functions and introduced ones are listed in Fig.~\ref{BFS-opt}. The first three functions in Fig.~\ref{BFS-opt} are search space constructor. They are newly introduced. \texttt{none} makes an empty search space. \texttt{unit} makes an space from one state. \texttt{step} makes a space from a thunk. The next function, \texttt{split}, is used to implement impure features (i.e. \texttt{ifte} and \texttt{once}). It takes a space \texttt{s-inf} and two continuations \texttt{ks} and \texttt{kf}. When \texttt{s-inf} contains some states, \texttt{ks} is called with the first state and the rest space. Otherwise, \texttt{kf} is called with no argument. Here `s' and `f' means `succeed' and `fail' respectively. The remaining functions do the same thing as before. Now it should be not difficult to see that \texttt{append-inf/fair} and \texttt{append-map-inf} do respect cost information.

Luckily, the change in \texttt{append-inf/fair} also fixes the miserable space inefficiency -- the use of \texttt{append} helps us to reuse the first bag of \texttt{t-inf}.

Noted that some functions in the list constitute a \emph{MonadPlus}: \texttt{none}, \texttt{unit}, \texttt{append-map-inf}, and \texttt{append-inf} correspond to \texttt{mzero}, \texttt{unit}, \texttt{bind}, and \texttt{mplus} respectively.

\begin{figure}
	\lstinputlisting{Figures/BFS-opt.rkt}	
	\caption{interface functions in optimized BFS}
	\label{BFS-opt}
\end{figure}

\subsection{comparison with Silvija's BFS}

In this section, we compare the pure subset of our optimized BFS with the BFS found in \citep{seres1999algebra}. We focus on the pure subset because Silvija's system is pure.

To compare efficiency, we translate her Haskell code into Racket (See supplements for the translated code), and embed it into miniKanren. The translation is fairly straightforward due to the similarity in both logic programming system and search space representation. The translated code is less efficient when running our benchmark. Details about efficiency difference are in section 6.

We prove the two search strategies are equivalent in Coq. Since search space can be infinite, we should use a co-inductive data type. However, Coq is too strict in the guardedness condition to accept a direct translation of the implementations. Therefore, we prove core theorems with finite search space instead. In order to generalize the conclusion to the cases with infinite search space, we prove a few more theorems saying that whenever we query answers lower than some finite cost, we can restrict goals to truncate search spaces at some finite depth without changing the query result. (See supplements for the formal proof)

;;TODO to be more specific?

@@

In this section, we compare the efficiency of search strategies. A concise description is in Table~\ref{compare-efficiency}. A hyphen means running out of memory. The first three benchmarks are taken from \citep{friedman_reasoned_2018}. Next two benchmarks about quine are modified from a similar test case in \citep{byrd2017unified}. The modifications are made to circumvent the need for symbolic constraint. Our version generates de Bruijnized expressions and forbids closures going into list. The two benchmarks differ in the \conde clause order of their relation interpreters. The last two benchmarks are about synthesizing expressions that evaluate to \texttt{'(I love you)}. This benchmark is also inspired by \citep{byrd2017unified}. Again, they differ in the \conde clause order of their relation interpreters. The first one has elimination rules (i.e. application, \texttt{car}, and \texttt{cdr}) at the end, while the other has them at the beginning. We conjecture that iDFS would perform badly in the second case because elimination rules complicate the problem when running backward. Our statistics support our conjecture.

In general, only iDFS and biDFS constantly perform well. Among them, biDFS seems to be less sensitive to change in \conde clause order (see the last four benchmarks). The other search strategies all have fair \disj{}. And they all perform badly in the \texttt{very-recursiveo} benchmark. However, the drawback of having fair \disj{} alone (i.e. fDFS) is not shown elsewhere. Fair \conj{} impose overhead constantly except in \texttt{appendo}. The reason might be that strategies with fair \conj{} tend to keep more intermediate states in the memory. Among the BFSs, our version performs better in most cases, and equally well elsewhere.

\begin{table}
\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline 
	benchmark & size & iDFS & biDFS & fDFS & optimized BFS & Silvija's BFS  \\
	\hline
	very-recursiveo & 100000 &  579 &  793 & 2131 & 1438 & 3617 \\
	                & 200000 & 1283 & 1610 & 3602 & 2803 & 4212 \\
	                & 300000 & 2160 & 2836 &    - & 6137 &    - \\
	\hline 
	appendo  & 100 &  31 &  41 &  42 &  31 &  68 \\ 
	         & 200 & 224 & 222 & 221 & 226 & 218 \\ 
	         & 300 & 617 & 634 & 593 & 631 & 622 \\ 
	\hline 
	reverseo & 10 &   5 &   3 &   3 &     38 &     85 \\ 
		     & 20 & 107 &  98 &  51 &   4862 &   5844 \\
		     & 30 & 446 & 442 & 485 & 123288 & 132159 \\ 
    \hline
    quine-1 & 1 &  71 &  44 & 69 & - & - \\ 
	        & 2 & 127 & 142 & 95 & - & - \\ 
	        & 3 & 114 & 114 & 93 & - & - \\ 
	\hline
    quine-2 & 1 & 147 & 112 &  56 & - & - \\ 
	        & 2 & 161 & 123 & 101 & - & - \\ 
	        & 3 & 289 & 189 & 104 & - & - \\ 
	\hline 
	'(I love you)-1 &  99 & 56 & 15 & 22 &  74 & 165 \\ 
	                & 198 & 53 & 72 & 55 &  47 &  74 \\
	                & 297 & 72 & 90 & 44 & 181 & 365 \\ 
    \hline
    '(I love you)-2 &  99 & 242 &  61 & 16 &  66 &  99 \\ 
	                & 198 & 445 & 110 & 60 &  42 &  64 \\
	                & 297 & 476 & 146 & 49 & 186 & 322 \\ 
	\hline 
\end{tabular}
\caption{The results of a quantitative evaluation: running times of benchmarks in milliseconds}
\label{compare-efficiency}
\end{table}

@@

Edward points out a disjunct complex would be `fair' if it is a full and balanced tree \citep{yang2010adventures}.

Silvija et al \citep{seres1999algebra} also describe a breadth-first search strategy. We proof their BFS is equivalent to ours. However, ours looks simpler and performs better in comparison with a straightforward translation of their Haskell code.

@@

%TODO contributions

% We devise a new search strategy, balanced interleaving DFS. The key idea is to make disjunct trees balanced. Changing the search strategy from iDFS to biDFS is not hard: 2 new functions and 1 modified macro.

% We also devise breadth-first search, whose intuition is similar to Seres's BFS. And we have proved their equivalence. We optimize our BFS with a queue.
